{\bf Landslide: Systematic Dynamic Race Detection in Kernel-space}

### 2012-05-09 defence/defence.tex
add fw/concl to outline
add more stuff

	{\bf Evaluation: Finding Races}
	\begin{itemize}
		\item Student user study
		\item Case study
	\end{itemize}
\end{frame}

@@

\begin{frame}{Finding Races in Kernel-Space} % slide: 15-410 6-week project why we need it in kernels
	15-410: Operating System Design and Implementation
	\begin{itemize}
		\item Project 3: students write a kernel in 6 weeks
		\item ``Pebbles'': UNIX-like, fully preemptible
	\end{itemize}
	\linegap

@@

	\linegap
	Context switcher
	\begin{itemize}
		\item Interleaving threads is no longer ``instantaneous''.
	\end{itemize}
	Runqueue tracking
	\begin{itemize}
		\item Which threads are runnable?
	\end{itemize}
	Thread lifecycle tracking
	\begin{itemize}
		\item When are threads created/destroyed?
	\end{itemize}
\end{frame}

@@

% slide: outline

@@

% slide: pebbles, 15-410, simics
\begin{frame}{Environment}
\end{frame}

@@

% slide: using -- annotations & in-landslide instrumentation
\begin{frame}{Using Landslide}
\end{frame}

% slide: POR -- mostly invisible, but (transition)
\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}

@@

\begin{frame}{Shared Memory Conflicts} % slide: ignoring shared memory accesses - venn diagram
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict0.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict1.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict2.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\textbf{Problem:} Many conflicts we don't care about
	\begin{itemize}
		\item Every transition accesses scheduler state
		\item Many transitions access ``unrelated'' global objects
	\end{itemize}
	\linegap

	{\bf Solution:} Only consider ``relevant'' memory conflicts
	\begin{itemize}
		\item Ignore scheduler, global objects ``known to be correct''
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: identifying bugs
\begin{frame}{Identifying Bugs}
	\textbf{How do we know we've found a bug?}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Limiting Decision Points}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

\subsection{Future Work}
\begin{frame}{Promising Directions}
	Top 15-410 students, effectively supervised, can find hard bugs faster
	\begin{itemize}
		\item Use as a debugging tool may help students learn more
	\end{itemize}
	\linegap

@@

\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic exploration in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour
		\item Relying on user's knowledge makes testing easier
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item After polishing, a useful tool for 15-410
		\item A first step towards sophisticated kernel debugging techniques
	\end{itemize}

>>>>

	{\bf Evaluation: Finding Races}
	\begin{itemize}
		\item Student user study
		\item Case study
	\end{itemize}
	\linegap

	{\bf Future Work and Conclusion}
\end{frame}

@@

\begin{frame}{Finding Races in Kernel-Space} % slide: 15-410 6-week project why we need it in kernels
	\textbf{15-410:} Operating System Design and Implementation
	\begin{itemize}
		\item Project 3: students write a kernel in 6 weeks.
		\item ``Pebbles'' is a UNIX-like, preemptible kernel specification.
	\end{itemize}
	\linegap

@@

	\linegap
	Context switching
	\begin{itemize}
		\item Non-deterministic timer-driven thread scheduling
	\end{itemize}
	Runqueue tracking
	\begin{itemize}
		\item Which threads are runnable?
	\end{itemize}
	Thread lifecycle tracking
	\begin{itemize}
		\item When are threads created/destroyed?
	\end{itemize}
\end{frame}

@@

% slide: outline
\begin{frame}{Outline}
\end{frame}

@@

% slide: pebbles, 15-410, simics
\begin{frame}{Environment}
	\textbf{Simics:} a full-system x86 simulator
	\begin{itemize}
		\item Students in 15-410 use Simics to test their kernels.
		\item Landslide is a transparently-embedded Simics module.
	\end{itemize}
\end{frame}

@@

% slide: using -- annotations & in-landslide instrumentation
\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

\begin{frame}{Decision Trace}
	\includegraphics[width=\textwidth]{found_a_bug.png}
\end{frame}

@@

\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}

\begin{frame}{Shared Memory Conflicts} % slide: ignoring shared memory accesses - venn diagram
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict0.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict1.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict2.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\textbf{Problem:} Many conflicts we don't care about
	\begin{itemize}
		\item Every transition accesses scheduler state.
		\item Many transitions access ``unrelated'' global objects.
	\end{itemize}
	\linegap

	{\bf Solution:} Only consider ``relevant'' memory conflicts.
	\begin{itemize}
		\item Ignore scheduler, global objects ``known to be correct''
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: identifying bugs
\begin{frame}{Identifying Bugs}
	\textbf{How do we know we've found a bug?}
	\linegap

	% TODO: does this slide want to be multicolumn, with an infinite loop diagram?
	Accurate conditions
	\begin{itemize}
		\item Kernel panic
		\item Use-after-free
		\item Deadlock
	\end{itemize}
	Uncertain conditions
	\begin{itemize}
		\item Memory leak
		\item Infinite loop
		\begin{itemize}
			\item Use structure of execution tree to judge stuckness.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Limiting Decision Points}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Outline}
\end{frame}

%\subsection{Future Work}
\begin{frame}{Promising Directions}
	Top 15-410 students, effectively supervised, can find hard bugs faster.
	\begin{itemize}
		\item Use as a debugging tool may help students learn more.
	\end{itemize}
	\linegap

@@

%\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic exploration in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour.
		\item Relying on user's knowledge makes testing easier.
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item After polishing, a useful tool for 15-410.
		\item A first step towards sophisticated kernel debugging techniques.
	\end{itemize}
